The promise of community-driven preprints in ecology and evolution

### 2024-05-03 ms/ms.qmd
Addeding a sentence about gender caveat

Research papers can take long periods of time to become published (see below). However, Early and Mid-Career Researchers (EMCRs) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to accelerate their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are one way EMCRs can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, we expect EMCRs to accept and adopt preprints more than colleagues at later career stages. At the same time, gender differences in preprint use and publication outcomes  have been observed in a number of other research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. Therefore, such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or women). To obtain an idea of an author's gender, we used the R package *gender* (vers. `r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We acknowledge that our approach does not capture self-assigned and non-binary genders. We only used algorithm-assigned names where the gender was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites. Our data on gender has only two missing values--one where the first name of the submitting author was missing and the other one for a collective submission.

>>>>

Research papers can take long periods of time to become published (see below). However, Early and Mid-Career Researchers (EMCRs) (~10 years post-PhD) are under pressure to publish rapidly to be competitive in job applications, promotions, and obtaining grants to accelerate their careers [@vale2015accelerating; @sarabipour2019value]. Preprints are one way EMCRs can achieve faster dissemination and greater visibility [@fu2019releasing]. As such, we expect EMCRs to accept and adopt preprints more than colleagues at later career stages. At the same time, gender differences in preprint use and publication outcomes have been observed in a number of other research fields, including ecology and evolutionary biology [@fox2019gender; @wehner2020comparison]. Therefore, such discrepancies are expected to manifest in preprint use on *EcoEvoRxiv*, but it is unclear to what extent. Understanding gender publishing patterns is challenging with observational data such as ours because we cannot know the gender of authors for certain but we can use a data-driven approach to ascertain the probability that a particular name is of a given gender (man or women). To obtain an idea of an author's gender, we used the R package *gender* (vers. `r utils::packageVersion("gender")`; @mullen2021predict) to predict the most likely gender of the submitting author of a preprint. We acknowledge that our approach does not capture self-assigned and non-binary genders. As such, our assumptions about an authors gender idendity could be incorrect. We only used algorithm-assigned names where the gender was identified with 95% certainty. For the remaining names, we performed manual searches to determine gender based on the pronouns and photographs from professional and personal websites. Our data on gender has only two missing values--one where the first name of the submitting author was missing and the other one for a collective submission.
